---
title: "ManyPrimates1_Phylogenetic_Models"
output:
  html_notebook:
    code_folding: hide
    css: style.css
    theme: paper
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(ape)
library(tidyverse)
library(readxl)
library(googledrive)
library(brms)
library(broom)
library(ggthemes)
library(tidybayes)
library(ggridges)
library(ggpubr)
library(parallel)
```

# Data processing

## Read in STM data file

```{r}
data <- read_csv("../data/merged_data/ManyPrimates_mp1_merged_data.csv")
```

## Read in species level predictors and scale them.

Species `hamlyns_monkey` and `allens_swamp_monkey` will be excluded from the phylogenetic analysis because we did not find data for some predictors for those species.

```{r}
# drive_download("https://docs.google.com/spreadsheets/d/1RCLvbNWFph6bHvupetx7RDiAu5BWXK6PumgRo44ZoeM/edit#gid=1909432990",
#                path = "../data/species_predictors.xlsx",
#                overwrite = T)

species_data <- read_xlsx("../data/species_predictors.xlsx", sheet = 1)%>%
  mutate_all(., .funs = tolower)%>%
  select(-contains("Source"), -contains("Comments"))%>%# remove columns with sources and comments
  rename(vocal_repertoire = `vocal_repertoire (# vocalization types)`,
         body_size = `body_size (kg)`)%>%
  mutate_if(is.character, tolower)%>%
  filter(species_english != "allens_swamp_monkey",
         species_english != "hamlyns_monkey") %>%
  mutate(vocal_repertoire = as.numeric(vocal_repertoire),
         life_expectancy = as.numeric(life_expectancy),
         group_size = as.numeric(group_size),
         home_range = as.numeric(home_range),
         feeding_budget = as.numeric(feeding_budget),
         day_journey_length= as.numeric(day_journey_length),
         resting_time_percent_activitybudget= as.numeric(resting_time_percent_activitybudget),
         feeding_budget= as.numeric(feeding_budget),
         dietary_breadth= as.numeric(dietary_breadth),
         percent_frugivory= as.numeric(percent_frugivory),
         body_size= as.numeric(body_size),
         home_range_body_site_ratio = home_range/body_size)%>%
  mutate(resting_time_percent_activitybudget = scale(resting_time_percent_activitybudget, center = T, scale = T),
         home_range_body_site_ratio = scale(home_range_body_site_ratio, center = T, scale = T),
         home_range = scale(home_range, center = T, scale = T),
         feeding_budget = scale(feeding_budget, center = T, scale = T),
         dietary_breadth = scale(dietary_breadth, center = T, scale = T),
         vocal_repertoire = scale(vocal_repertoire, center = T, scale = T),
         group_size = scale(group_size, center = T, scale = T),
         day_journey_length = scale(day_journey_length, center = T, scale = T),
         percent_frugivory = scale(percent_frugivory, center = T, scale = T),
         feeding_budget = scale(feeding_budget, center = T, scale =T))

```

## Merge experimental data with species-level predictors

Exclude species for which we have no species-level data. `Gorilla` has to be re-coded to match the species name in the species-level predictor file.

```{r}
phylo_data <- data %>%
  rename(species_english = species)%>% # create a common column to merge by
  filter(species_english != "allens_swamp_monkey",
         species_english != "hamlyns_monkey") %>%
  left_join(species_data%>%select(-species_latin,-family))

# check if there are remaining NAs for the predictors

phylo_data %>%
  filter_at(vars(21:32),any_vars(is.na(.)))
```

## Recode the "short/long/medium" condition variable into medium-centred numeric delays:

```{r}
phylo_data <- phylo_data %>%
  mutate(delay = if_else(condition == "short", -1,
                         if_else(condition == "medium", 0, 1)),
         trial = scale(trial, center = T, scale = T))
```

## Save the merged/recoded data for use in subsequent Rmds

```{r}
write_rds(phylo_data, "../saves/phylo_data.rds")
```

# Phylogenetic tree

Tree-related processing:

 * Read phylogenetic tree from file
 * Check that all species labels in the data also occur as tip labels in the tree
 * Prune any species not in the data from the tree
 * Compute covariance matrix from tree under Brownian motion model

```{r}
tree <- read.tree("../phylo/mp_species.tree")
data_species <- phylo_data$species_english %>% unique
tree_species <- tree$tip.label
stopifnot(length(setdiff(data_species, intersect(tree_species, data_species))) == 0)
tree <- keep.tip(tree, data_species)
A <- vcv.phylo(tree)
```

# Phylogenetic models

```{r}
formulas <- list()
priors <- list()
```

## Model 0: Phylogenetic baseline model

```{r}
model_name <- "base_model"
formulas[[model_name]] <- correct ~ delay + (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(skew_normal(0, 1.5, 3), class=Intercept) +
                        prior(skew_normal(-1.4, 2, 3), coef=delay) +
                        prior(normal(0, 1), class=sd, coef=Intercept, group=subject_site) +
                        prior(normal(0, 1), class=sd, coef=Intercept, group=site) +
                        prior(normal(0, 1), class=sd, coef=Intercept, group=species_english) +
                        prior(normal(0, 0.5), class=sd, coef=delay, group=subject_site) +
                        prior(normal(0, 0.5), class=sd, coef=delay, group=site) +
                        prior(normal(0, 0.5), class=sd, coef=delay, group=species_english) +
                        prior(normal(0, 0.5), class=sd, coef=trial, group=subject_site) +
                        prior(normal(0, 0.5), class=sd, coef=trial, group=site) +
                        prior(normal(0, 0.5), class=sd, coef=trial, group=species_english) +
                        prior("lkj(1)", class = "cor")
```

## Model 1: Diurnal resting-time

Numeric predictor: `resting_time_percent_activitybudget`.

```{r}
model_name <- "diurnal_resting_model"
formulas[[model_name]] <- correct ~ delay * resting_time_percent_activitybudget +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=resting_time_percent_activitybudget) +
                        prior(normal(0, 1), coef=`delay:resting_time_percent_activitybudget`)
```

## Model 2: Color vision

Categorical predictor: `color_vision`. 

```{r}
model_name <- "color_vision_model"
formulas[[model_name]] <- correct ~ delay * color_vision +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 2), coef=color_visionpolymorphic) +
                        prior(normal(0, 2), coef=color_visiontrichromatic) +
                        prior(normal(0, 2), coef=`delay:color_visionpolymorphic`) +
                        prior(normal(0, 2), coef=`delay:color_visiontrichromatic`)
```

## Model 3: Home_range/body size ratio, time spent feeding and range of consumed food items

numerical predictors: `home_range_body_site_ratio`, `feeding_budget` and `dietary_breadth`.

```{r}
model_name <- "home_range_feeding_budget_diet_breath_model"
formulas[[model_name]] <- correct ~ delay * home_range_body_site_ratio + delay * feeding_budget * dietary_breadth +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=home_range_body_site_ratio) +
                        prior(normal(0, 1), coef=`delay:dietary_breadth`) +
                        prior(normal(0, 1), coef=`delay:home_range_body_site_ratio`) +
                        prior(normal(0, 1), coef=`delay:feeding_budget`) +
                        prior(normal(0, 1), coef=`delay:feeding_budget:dietary_breadth`) +
                        prior(normal(0, 1), coef=dietary_breadth) +
                        prior(normal(0, 1), coef=feeding_budget) +
                        prior(normal(0, 1), coef=`feeding_budget:dietary_breadth`)
```

## Model 3a: Home range

numerical predictor: `home_range`.

```{r}
model_name <- "home_range_model"
formulas[[model_name]] <- correct ~ delay * home_range +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=home_range) +
                        prior(normal(0, 1), coef=`delay:home_range`)
```

## Model 3b: Feeding budget

numerical predictor: `feeding_budget`

```{r}
model_name <- "feeding_budget_model"
formulas[[model_name]] <- correct ~ delay * feeding_budget +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=feeding_budget) +
                        prior(normal(0, 1), coef=`delay:feeding_budget`)
```

## Model 4: Vocal repertoire

Numeric predictor: `vocal_repertoire`. 

```{r}
model_name <- "vocal_repertoire_model"
formulas[[model_name]] <- correct ~ delay * vocal_repertoire +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=vocal_repertoire) +
                        prior(normal(0, 1), coef=`delay:vocal_repertoire`)
```

## Model 5: Dietary breath

Numeric predictor: `dietary_breadth`. 

```{r}
model_name <- "dietary_breadth_model"
formulas[[model_name]] <- correct ~ delay * dietary_breadth +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=dietary_breadth) +
                        prior(normal(0, 1), coef=`delay:dietary_breadth`)
```

## Model 6: Group size

Numeric predictor: `group_size`. 

```{r}
model_name <- "group_size_model"
formulas[[model_name]] <- correct ~ delay * group_size +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=group_size) +
                        prior(normal(0, 1), coef=`delay:group_size`)
```

## Model 7: Day journey length and group size

Numeric predictors: `day_journey_length` and `group_size`. 

```{r}
model_name <- "day_journey_group_size_model"
formulas[[model_name]] <- correct ~ delay * day_journey_length * group_size +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=group_size) +
                        prior(normal(0, 1), coef=day_journey_length) +
                        prior(normal(0, 1), coef=`day_journey_length:group_size`) +
                        prior(normal(0, 1), coef=`delay:group_size`) +
                        prior(normal(0, 1), coef=`delay:day_journey_length`) +
                        prior(normal(0, 1), coef=`delay:day_journey_length:group_size`)
```

## Model 7a: Day journey length

Numeric predictor: `day_journey_length`.

```{r}
model_name <- "day_journey_model"
formulas[[model_name]] <- correct ~ delay * day_journey_length +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=day_journey_length) +
                        prior(normal(0, 1), coef=`delay:day_journey_length`)
```

## Model 8: Percent frugivory and terrestriality

Numeric predictor: `percent_frugivory`. Categorical predictor and `terrestriality`. 

```{r}
model_name <- "frugivory_terrestriality_model"
formulas[[model_name]] <- correct ~ delay * percent_frugivory * terrestriality +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 2), coef=terrestrialityterrestrial) +
                        prior(normal(0, 1), coef=percent_frugivory) +
                        prior(normal(0, 2), coef=`percent_frugivory:terrestrialityterrestrial`) +
                        prior(normal(0, 2), coef=`delay:terrestrialityterrestrial`) +
                        prior(normal(0, 1), coef=`delay:percent_frugivory`) +
                        prior(normal(0, 2), coef=`delay:percent_frugivory:terrestrialityterrestrial`)
```

## Model 8a: Percent frugivory

Numeric predictor: `percent_frugivory`.

```{r}
model_name <- "frugivory_model"
formulas[[model_name]] <- correct ~ delay * percent_frugivory +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=percent_frugivory) +
                        prior(normal(0, 1), coef=`delay:percent_frugivory`)
```

## Model 8b: Terrestriality

Categorical predictor and `terrestriality`. 

```{r}
model_name <- "terrestriality_model"
formulas[[model_name]] <- correct ~ delay * terrestriality +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 2), coef=terrestrialityterrestrial) +
                        prior(normal(0, 2), coef=`delay:terrestrialityterrestrial`)
```

## Model 9: Diet diversity

Categorical predictor: `diet_diversity`. 

```{r}
model_name <- "diet_diversity_model"
formulas[[model_name]] <- correct ~ delay * diet_diversity +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 2), coef=diet_diversityfolivore_frugivore) +
                        prior(normal(0, 2), coef=diet_diversityfrugivore) +
                        prior(normal(0, 2), coef=diet_diversitygummivore) +
                        prior(normal(0, 2), coef=diet_diversityinsectivore_frugivore) +
                        prior(normal(0, 2), coef=diet_diversityomnivore) +
                        prior(normal(0, 2), coef=`delay:diet_diversityfolivore_frugivore`) +
                        prior(normal(0, 2), coef=`delay:diet_diversityfrugivore`) +
                        prior(normal(0, 2), coef=`delay:diet_diversitygummivore`) +
                        prior(normal(0, 2), coef=`delay:diet_diversityinsectivore_frugivore`) +
                        prior(normal(0, 2), coef=`delay:diet_diversityomnivore`)
```

## Run all models in parallel

Models are saved to `saves/` folder.

```{r}

# First, add the common base prior to all other priors
stopifnot(names(formulas)[1] == "base_model")
for(i in 2:length(priors)) {
  priors[[i]] <- priors[[1]] + priors[[i]]
}

# Derive filenames for model results from model names
filenames <- paste("../saves/phylo_", names(formulas), ".rds", sep="")

# Define a function to run, process and save a model given a formula and prior
run_model <- function(formula, prior, chains, iter, filename) {
  print(c(filename, formula, prior))
  model <- brm(formula, family=bernoulli(),
               data=phylo_data,
               data2 = list(A = A),
               prior=prior,
               chains=chains,
               iter=iter,
               cores=chains,
            )
  model <- add_criterion(model, c("loo", "waic"))
  saveRDS(model, filename)
}

# Run and save all models in parallel
chains <- 8
iter <- 5000
available_cores <- detectCores()
simultaneous_models <- floor(available_cores / chains)

## Turn warnings into errors so we won't get past the call to mcmapply if
## something goes wrong
options(warn=2)

mcmapply(run_model, formulas, priors, chains, iter, filenames,
         mc.cores=simultaneous_models)

## Restore default warning behaviour
options(warn=0)

# Load all model result and assign them nice variables names ready for the
# model comparison code below
for(i in 1:length(formulas)) {
  model <- readRDS(filenames[i])
  assign(names(formulas)[i], model)
}
```

# Appendix: Confirmatory models

Run a model that does not include the phylogenetic tree in the random effects. This will give us an additional angle on the phylogenetic signal in the data. 

## Model 10: No phylogeny model

```{r}
no_phylogeny_model <- brm(correct ~ delay +
                   (1 + delay + trial | subject_site + site + species_english),
                   data=phylo_data,
                   family=bernoulli(),
                   chains = 8,
                   iter= 5000,
                   cores= 8,
                   prior = prior(skew_normal(0, 1.5, 3), class=Intercept) +
                           prior(skew_normal(-1.4, 2, 3), class=b, coef=delay) +
                           prior(normal(0, 1), class=sd, coef=Intercept, group=subject_site) +
                           prior(normal(0, 1), class=sd, coef=Intercept, group=site) +
                           prior(normal(0, 1), class=sd, coef=Intercept, group=species_english) +
                           prior(normal(0, 0.5), class=sd, coef=delay, group=subject_site) +
                           prior(normal(0, 0.5), class=sd, coef=delay, group=site) +
                           prior(normal(0, 0.5), class=sd, coef=delay, group=species_english) +
                           prior(normal(0, 0.5), class=sd, coef=trial, group=subject_site) +
                           prior(normal(0, 0.5), class=sd, coef=trial, group=site) +
                           prior(normal(0, 0.5), class=sd, coef=trial, group=species_english) +
                           prior("lkj(1)", class = "cor")
                )

no_phylogeny_model <- add_criterion(no_phylogeny_model, c("loo", "waic"))

no_phylogeny_model %>% saveRDS("../saves/phylo_no_phylogeny_model.rds")
```

