---
title: "ManyPrimates1_Phylogenetic_Model_Comparison"
output:
  html_notebook:
    code_folding: hide
    css: style.css
    theme: paper
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(ape)
library(tidyverse)
library(readxl)
library(googledrive)
library(brms)
library(broom)
library(ggthemes)
library(tidybayes)
library(ggridges)
library(ggpubr)
library(parallel)
```


# NOTE
Running the models takes a long time (12h on a HPC) - outputs are too large to put on GitHub (~7GB). If you want to access the model outputs directly, you have to download them via the following link. Code below expects files to be in `saves/`.

If the code doesn't work, you can download the files here: `https://keeper.mpdl.mpg.de/f/e78a9963b58f4a9190d8/?dl=1` but you will have to put them manually into the `saves` directory.

```{r}
# download zip file from Max Planck cloud storage
options(timeout=3600) # 1h, maybe you need more depending on your internet speed
download.file("https://keeper.mpdl.mpg.de/f/e78a9963b58f4a9190d8/?dl=1", 
              destfile = "../saves/saves.zip")

# unpack zip file to get .rds files
unzip("../saves/saves.zip", exdir = "../saves/")
```



# Data processing

## Read in STM data file

```{r}
data <- read_csv("../data/merged_data/ManyPrimates_mp1_merged_data.csv")
```

# Model comparison

## Read in models

```{r}
base_model  <- readRDS("../saves/phylo_base_model.rds")
diurnal_resting_model <- readRDS("../saves/phylo_diurnal_resting_model.rds")
color_vision_model <- readRDS("../saves/phylo_color_vision_model.rds")
home_range_feeding_budget_diet_breath_model <- readRDS("../saves/phylo_home_range_feeding_budget_diet_breath_model.rds")
vocal_repertoire_model <- readRDS("../saves/phylo_vocal_repertoire_model.rds")
dietary_breadth_model <- readRDS("../saves/phylo_dietary_breadth_model.rds")
group_size_model <- readRDS("../saves/phylo_group_size_model.rds")
day_journey_group_size_model <- readRDS("../saves/phylo_day_journey_group_size_model.rds")
frugivory_terrestriality_model <- readRDS("../saves/phylo_frugivory_terrestriality_model.rds")
diet_diversity_model <- readRDS("../saves/phylo_diet_diversity_model.rds")
home_range_model <- readRDS("../saves/phylo_home_range_model.rds")
feeding_budget_model <- readRDS("../saves/phylo_feeding_budget_model.rds")
day_journey_model <- readRDS("../saves/phylo_day_journey_model.rds")
frugivory_model <- readRDS("../saves/phylo_frugivory_model.rds")
terrestriality_model <- readRDS("../saves/phylo_terrestriality_model.rds")
```

## Models from modelling challenge

```{r}
# Compute WAIC scores and SE
comp <-  loo_compare(
  base_model, 
  diurnal_resting_model,
  color_vision_model,
  home_range_feeding_budget_diet_breath_model,
  vocal_repertoire_model,
  dietary_breadth_model,
  group_size_model,
  day_journey_group_size_model,
  frugivory_terrestriality_model,
  diet_diversity_model,
  criterion = "waic")%>%
  as_tibble(rownames = "model")
  
# Compute WAIC weights
weights <- model_weights(
  base_model, 
  diurnal_resting_model,
  color_vision_model,
  home_range_feeding_budget_diet_breath_model,
  vocal_repertoire_model,
  dietary_breadth_model,
  group_size_model,
  day_journey_group_size_model,
  frugivory_terrestriality_model,
  diet_diversity_model, weights = "waic")%>%
  as_tibble() %>% 
  rename(weight = value) %>% 
  mutate(model  = c("base_model", 
                    "diurnal_resting_model",
                    "color_vision_model",
                    "home_range_feeding_budget_diet_breath_model",
                    "vocal_repertoire_model",
                    "dietary_breadth_model",
                    "group_size_model",
                    "day_journey_group_size_model",
                    "frugivory_terrestriality_model",
                    "diet_diversity_model"),
         weight = weight %>% round(digits = 2)) %>% 
  select(model, weight) 


# Combine weights and scores
model_comparison <- comp %>%
  select(model,waic, se_waic)%>%
  left_join(weights)%>%
  arrange(desc(weight))

#write_csv(model_comparison, "../saves/model_comparison.csv")

model_comparison <- read_csv("../saves/model_comparison.csv")

model_comparison
```
## Exploratory analysis:Include main effects models for all predictors

Some models were submitted with interaction terms between variables. The additional models below were built so that each predictor that was submitted as part of an interaction is also included in a model with only that predictor. 

```{r}
# Compute WAIC scores and SE
comp_exp <-  loo_compare(
  base_model,    # models from modelling challnege
  diurnal_resting_model,
  color_vision_model,
  home_range_feeding_budget_diet_breath_model,
  vocal_repertoire_model,
  dietary_breadth_model,
  group_size_model,
  day_journey_group_size_model,
  frugivory_terrestriality_model,
  diet_diversity_model,
  
  home_range_model,      # main effects models 
  feeding_budget_model,
  day_journey_model,
  frugivory_model,
  terrestriality_model,
  
  criterion = "waic")%>%
  as_tibble(rownames = "model")
  
# Compute WAIC weights
weights_exp <- model_weights(
  base_model, 
  diurnal_resting_model,
  color_vision_model,
  home_range_feeding_budget_diet_breath_model,
  vocal_repertoire_model,
  dietary_breadth_model,
  group_size_model,
  day_journey_group_size_model,
  frugivory_terrestriality_model,
  diet_diversity_model, 
  home_range_model,
  feeding_budget_model,
  day_journey_model,
  frugivory_model,
  terrestriality_model,
  weights = "waic")%>%
  as_tibble() %>% 
  rename(weight = value) %>% 
  mutate(model  = c("base_model", 
                    "diurnal_resting_model",
                    "color_vision_model",
                    "home_range_feeding_budget_diet_breath_model",
                    "vocal_repertoire_model",
                    "dietary_breadth_model",
                    "group_size_model",
                    "day_journey_group_size_model",
                    "frugivory_terrestriality_model",
                    "diet_diversity_model",
                    "home_range_model",
                    "feeding_budget_model",
                    "day_journey_model",
                    "frugivory_model",
                    "terrestriality_model"),
         weight = weight %>% round(digits = 2)) %>% 
  select(model, weight) 


# Combine weights and scores
model_comparison_exp <- comp_exp %>%
  select(model,waic, se_waic)%>%
  left_join(weights_exp)%>%
  arrange(desc(weight))

#write_csv(model_comparison_exp, "../saves/model_comparison_exp.csv")

model_comparison_exp <- read_csv("../saves/model_comparison_exp.csv")

model_comparison_exp
```
# Visualize winning models and baseline

## Posterior distributions for predictors
```{r}
model_draws <-  bind_rows(
  vocal_repertoire_model%>% posterior_samples(pars = c("b_delay","b_vocal_repertoire","b_delay:vocal_repertoire"))%>%
    rename(delay = b_delay,
           predictor = b_vocal_repertoire,
           interaction = `b_delay:vocal_repertoire`)%>%
    mutate(model= "vocal_repertoire"),
  base_model%>% posterior_samples(pars = c("b_delay"))%>%
    rename(delay = b_delay)%>%
    mutate(model= "baseline")
)%>%
  pivot_longer(names_to = "estimate", values_to = "value", cols = c(delay,predictor,interaction ))

model_draws %>%
  ggplot(.,aes(x = value, y= model, fill = factor(stat(quantile))))+
  geom_vline(xintercept = 0, lty = 2)+
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = c(0.025, 0.975)
  ) +
  facet_grid(~estimate)+
  theme_minimal() +
  scale_fill_manual(
   values = c("firebrick", "NA", "firebrick")
  )+
  guides(fill = F)
```
##Correlation between predictor and stm performance

Performance estimate is the species intercept from the baseline model. Not sure if that makes sense.

```{r}
base_model%>% spread_draws(r_species_english[species_english,term])%>%
  filter(term == "Intercept")%>%
  rename(stm_estimate = "r_species_english")%>%
  group_by(species_english)%>%
  median_qi(stm_estimate)%>%
  left_join(species_data%>%
              select(species_english,superordinate_group,vocal_repertoire,dietary_breadth))%>%
  pivot_longer(cols = c("vocal_repertoire","dietary_breadth"), names_to = "variable", values_to = "predictor")%>%
  ggplot(aes(x = predictor, y = stm_estimate, col = superordinate_group))+
  facet_grid(~variable)+
  geom_point(alpha = .5)+
  geom_smooth(method = lm, se = F, lty = 2, col = "grey")+
  stat_cor(method = "pearson", inherit.aes = F, aes(x = predictor, y = stm_estimate), r.accuracy = 0.01, cor.coef.name = "r")+
  theme_minimal()
  
```


