---
title: "ManyPrimates1_Phylogenetic_Models_Prior_Sensitivity"
output:
  html_notebook:
    code_folding: hide
    css: style.css
    theme: paper
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(ape)
library(tidyverse)
library(readxl)
library(googledrive)
library(brms)
library(broom)
library(ggthemes)
library(tidybayes)
library(ggridges)
library(ggpubr)
library(parallel)
```

# NOTE
Running the models takes a long time - outputs are too large to put on GitHub (~6GB). If you want to access the model outputs directly, you have to download them via the following link. Code that follows will put them into the `saves/` folder.

```{r}
# download zip file from Max Planck cloud storage
options(timeout=3600) # 1h, maybe you need more depending on your internet speed
download.file("https://keeper.mpdl.mpg.de/f/f7b96fe82bd34a0caa74/?dl=1", 
              destfile = "../saves/saves.zip")
# unpack zip file to get .rds files
unzip("../saves/saves.zip", exdir = "../saves/")
```

If the code doesn't work, you can download the files here: `https://keeper.mpdl.mpg.de/f/f7b96fe82bd34a0caa74/?dl=1`.

# Data processing

## Read in the already merged and recoded data and pruned tree from the
## earlier run of the phylogenetic models


```{r}
phylo_data <- read_rds("../saves/phylo_data.rds")
tree <- read_rds("../saves/pruned_tree.rds")
A <- vcv.phylo(tree)
```

# Phylogenetic models

```{r}
formulas <- list()
priors <- list()
```

## Model 0: Phylogenetic baseline model

```{r}
model_name <- "base_model"
formulas[[model_name]] <- correct ~ delay + (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(skew_normal(0, 1.5, 3), class=Intercept) +
                        prior(skew_normal(-1.4, 2, -3), coef=delay1) +
                        prior(normal(0, 1), class=sd, coef=Intercept, group=subject_site) +
                        prior(normal(0, 1), class=sd, coef=Intercept, group=site) +
                        prior(normal(0, 0.5), class=sd, coef=delay1, group=subject_site) +
                        prior(normal(0, 0.5), class=sd, coef=delay1, group=site) +
                        prior(normal(0, 0.5), class=sd, coef=trial, group=subject_site) +
                        prior(normal(0, 0.5), class=sd, coef=trial, group=site) +
                        prior("lkj(1)", class = "cor")

reference_species_priors <- prior(normal(0, 1), class=sd, coef=Intercept, group=species_english) +
                        prior(normal(0, 0.5), class=sd, coef=delay1, group=species_english) +
                        prior(normal(0, 0.5), class=sd, coef=trial, group=species_english)


weaker_species_priors <- prior(normal(0, 0.5), class=sd, coef=Intercept, group=species_english) +
                         prior(normal(0, 0.25), class=sd, coef=delay1, group=species_english) +
                         prior(normal(0, 0.25), class=sd, coef=trial, group=species_english)

weakest_species_priors <- prior(normal(0, 0.25), class=sd, coef=Intercept, group=species_english) +
                          prior(normal(0, 0.125), class=sd, coef=delay1, group=species_english) +
                          prior(normal(0, 0.125), class=sd, coef=trial, group=species_english)

stronger_species_priors <- prior(normal(0, 2.0), class=sd, coef=Intercept, group=species_english) +
                           prior(normal(0, 1.0), class=sd, coef=delay1, group=species_english) +
                           prior(normal(0, 1.0), class=sd, coef=trial, group=species_english)
```

## Model 3a: Home range

numerical predictor: `home_range`.

```{r}
model_name <- "home_range_model"
formulas[[model_name]] <- correct ~ delay * home_range +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=home_range) +
                        prior(normal(0, 1), coef=`delay1:home_range`)
```

## Model 4: Vocal repertoire

Numeric predictor: `vocal_repertoire`. 

```{r}
model_name <- "vocal_repertoire_model"
formulas[[model_name]] <- correct ~ delay * vocal_repertoire +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=vocal_repertoire) +
                        prior(normal(0, 1), coef=`delay1:vocal_repertoire`)
```

## Model 5: Dietary breath

Numeric predictor: `dietary_breadth`. 

```{r}
model_name <- "dietary_breadth_model"
formulas[[model_name]] <- correct ~ delay * dietary_breadth +
                          (1 + delay + trial | subject_site + site + gr(species_english, cov=A))
priors[[model_name]] <- prior(normal(0, 1), coef=dietary_breadth) +
                        prior(normal(0, 1), coef=`delay1:dietary_breadth`)
```

## Run all models in parallel

Models are saved to `saves/` folder.

```{r}

# Weaker phylogeny
reference_priors <- list()
weaker_priors <- list()
weakest_priors <- list()
stronger_priors <- list()
reference_priors[[1]] <- priors[[1]] + reference_species_priors
weaker_priors[[1]] <- priors[[1]] + weaker_species_priors
weakest_priors[[1]] <- priors[[1]] + weakest_species_priors
stronger_priors[[1]] <- priors[[1]] + stronger_species_priors
for(i in 2:length(priors)) {
  reference_priors[[i]] <- reference_priors[[1]] + priors[[i]]
  weaker_priors[[i]] <- weaker_priors[[1]] + priors[[i]]
  weakest_priors[[i]] <- weakest_priors[[1]] + priors[[i]]
  stronger_priors[[i]] <- stronger_priors[[1]] + priors[[i]]
}
stopifnot(length(reference_priors) == length(formulas))
stopifnot(length(weaker_priors) == length(formulas))
stopifnot(length(weakest_priors) == length(formulas))
stopifnot(length(stronger_priors) == length(formulas))

# Define a function to run, process and save a model given a formula and prior
run_model <- function(formula, prior, chains, iter, filename) {
  print(c(filename, formula, prior))
  model <- brm(formula, family=bernoulli(),
               data=phylo_data,
               data2 = list(A = A),
               prior=prior,
               chains=chains,
               iter=iter,
               cores=chains,
            )
  model <- add_criterion(model, c("loo", "waic"))
  saveRDS(model, filename)
}

# Run and save all models in parallel
chains <- 8
iter <- 5000
available_cores <- detectCores()
simultaneous_models <- floor(available_cores / chains)

## Turn warnings into errors so we won't get past the call to mcmapply if
## something goes wrong
#options(warn=2)

# Weaker
filenames <- paste("../saves/weaker_phylo_", names(formulas), ".rds", sep="")
mcmapply(run_model, formulas, weaker_priors, chains, iter, filenames,
         mc.cores=simultaneous_models)

# Weakest
filenames <- paste("../saves/weakest_phylo_", names(formulas), ".rds", sep="")
mcmapply(run_model, formulas, weakest_priors, chains, iter, filenames,
         mc.cores=simultaneous_models)

# Stronger
filenames <- paste("../saves/stronger_phylo_", names(formulas), ".rds", sep="")
mcmapply(run_model, formulas, stronger_priors, chains, iter, filenames,
         mc.cores=simultaneous_models)

## Restore default warning behaviour
#options(warn=0)

```


```{r}
# Do and save comparisons
weaker_base_model <- readRDS("../saves/weaker_phylo_base_model.rds")
weakest_base_model <- readRDS("../saves/weakest_phylo_base_model.rds")
stronger_base_model <- readRDS("../saves/stronger_phylo_base_model.rds")


model_names <- c("base_model", "vocal_repertoire_model", "dietary_breadth_model", "home_range_model")

# weaker prior
weaker_vocal_repertoire_model <- readRDS("../saves/weaker_phylo_vocal_repertoire_model.rds")
weaker_dietary_breadth_model <- readRDS("../saves/weaker_phylo_dietary_breadth_model.rds")
weaker_home_range_model <- readRDS("../saves/weaker_phylo_home_range_model.rds")
weaker_comp <- loo_compare(weaker_base_model, weaker_vocal_repertoire_model,
			     weaker_dietary_breadth_model, weaker_home_range_model,
                             criterion = "waic") %>% as_tibble %>% mutate(model=model_names)
write_csv(as_tibble(weaker_comp), "../saves/weaker_phylogenetic_model_comparison.csv")

# weakes prior
weakest_vocal_repertoire_model <- readRDS("../saves/weakest_phylo_vocal_repertoire_model.rds")
weakest_dietary_breadth_model <- readRDS("../saves/weakest_phylo_dietary_breadth_model.rds")
weakest_home_range_model <- readRDS("../saves/weakest_phylo_home_range_model.rds")
weakest_comp <- loo_compare(weakest_base_model, weakest_vocal_repertoire_model,
			     weakest_dietary_breadth_model, weakest_home_range_model,
                             criterion="waic") %>% as_tibble %>% mutate(model=model_names)
write_csv(as_tibble(weakest_comp), "../saves/weakest_phylogenetic_model_comparison.csv")

# stronger prior
stronger_vocal_repertoire_model <- readRDS("../saves/stronger_phylo_vocal_repertoire_model.rds")
stronger_dietary_breadth_model <- readRDS("../saves/stronger_phylo_dietary_breadth_model.rds")
stronger_home_range_model <- readRDS("../saves/stronger_phylo_home_range_model.rds")
stronger_comp <- loo_compare(stronger_base_model, stronger_vocal_repertoire_model,
			     stronger_dietary_breadth_model, stronger_home_range_model,
                             criterion="waic") %>% as_tibble %>% mutate(model=model_names)
write_csv(as_tibble(stronger_comp), "../saves/stronger_phylogenetic_model_comparison.csv")
```

