---
title: "ManyPrimates1_Confirmatory_Analysis"
output:
  html_notebook:
    code_folding: hide
    css: style.css
    theme: paper
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(ape)
library(tidyverse)
library(readxl)
library(googledrive)
library(brms)
library(broom)
library(ggthemes)
library(tidybayes)
library(ggridges)
library(ggpubr)
```


# NOTE
Running the models takes a long time (12h on a HPC) - outputs are too large to put on GitHub (~7GB). If you want to access the model outputs directly, you have to download them via the following link. Code below expects files to be in `saves/`.

If the code doesn't work, you can download the files here: `https://keeper.mpdl.mpg.de/f/e78a9963b58f4a9190d8/?dl=1` but you will have to put them manually into the `saves` directory.

# Data processing

## Read in STM data file

Read in data, re-code condition and scale predictors. 

```{r}
data <- read_csv("../data/merged_data/ManyPrimates_mp1_merged_data.csv")%>%
  mutate(delay = if_else(condition == "short", -1,
                         if_else(condition == "medium", 0, 1)),
         trial = scale(trial, center = T, scale = T),
         norm_age = scale(norm_age, center = T, scale = T),
         cup_distance = scale(cup_distance, center = T, scale = T),
         board_size = scale(board_size, center = T, scale = T))
```

# Confirmatory analysis

## Full model

```{r}
full_model <- brm(correct ~ delay +
                    task_experience + norm_age + cup_distance + board_size + trial +
                   (1 + delay + trial | subject_site + site + species),
                   data=data,
                   family=bernoulli(),
                   chains = 4,
                   iter= 5000,
                   cores= 4,
                   prior = prior(skew_normal(0, 1.5, 3), class=Intercept) +
                           prior(skew_normal(-1.4, 2, 3), class=b, coef=delay) +
                           prior(normal(0, 1), class=b, coef=board_size) +
                           prior(normal(0, 1), class=b, coef=trial) +
                           prior(normal(0, 1), class=b, coef=cup_distance) +
                           prior(normal(0, 1), class=b, coef=norm_age) +
                           prior(normal(0, 2), class=b, coef=task_experienceyes) +
                           prior(normal(0, 1), class=sd, coef=Intercept, group=subject_site) +
                           prior(normal(0, 1), class=sd, coef=Intercept, group=site) +
                           prior(normal(0, 1), class=sd, coef=Intercept, group=species) +
                           prior(normal(0, 0.5), class=sd, coef=delay, group=subject_site) +
                           prior(normal(0, 0.5), class=sd, coef=delay, group=site) +
                           prior(normal(0, 0.5), class=sd, coef=delay, group=species) +
                           prior(normal(0, 0.5), class=sd, coef=trial, group=subject_site) +
                           prior(normal(0, 0.5), class=sd, coef=trial, group=site) +
                           prior(normal(0, 0.5), class=sd, coef=trial, group=species) +
                           prior("lkj(1)", class = "cor")
                )

summary(full_model)

full_model <- add_criterion(full_model, "waic")

full_model %>% saveRDS("../saves/full_model.rds")

#full_model <- readRDS("../saves/full_model.rds")
```

## Null model

```{r}
null_model <- brm(correct ~ task_experience + norm_age + cup_distance + board_size + trial +
                   (1 + delay + trial | subject_site + site + species),
                   data=data,
                   family=bernoulli(),
                   chains = 4,
                   iter= 5000,
                   cores= 4,
                   prior = prior(skew_normal(0, 1.5, 3), class=Intercept) +
                           #prior(skew_normal(-1.4, 2, 3), class=b, coef=delay) +
                           prior(normal(0, 1), class=b, coef=board_size) +
                           prior(normal(0, 1), class=b, coef=trial) +
                           prior(normal(0, 1), class=b, coef=cup_distance) +
                           prior(normal(0, 1), class=b, coef=norm_age) +
                           prior(normal(0, 2), class=b, coef=task_experienceyes) +
                           prior(normal(0, 1), class=sd, coef=Intercept, group=subject_site) +
                           prior(normal(0, 1), class=sd, coef=Intercept, group=site) +
                           prior(normal(0, 1), class=sd, coef=Intercept, group=species) +
                           prior(normal(0, 0.5), class=sd, coef=delay, group=subject_site) +
                           prior(normal(0, 0.5), class=sd, coef=delay, group=site) +
                           prior(normal(0, 0.5), class=sd, coef=delay, group=species) +
                           prior(normal(0, 0.5), class=sd, coef=trial, group=subject_site) +
                           prior(normal(0, 0.5), class=sd, coef=trial, group=site) +
                           prior(normal(0, 0.5), class=sd, coef=trial, group=species) +
                           prior("lkj(1)", class = "cor")
                )

summary(null_model)

null_model <- add_criterion(null_model, "waic")

null_model %>% saveRDS("../saves/null_model.rds")

#null_model <- readRDS("../saves/null_model.rds")

```

## Model comparison

```{r}
comp <-  loo_compare(
  full_model, 
  null_model,
  criterion = "waic")%>%
  as_tibble(rownames = "model")

weights <- model_weights(full_model, null_model, weights = "waic")%>%
  as_tibble() %>% 
  rename(weight = value) %>% 
  mutate(model  = c("full_model", "null_model"),
         weight = weight %>% round(digits = 4)) %>% 
  select(model, weight) 

# Combine weights and scores
model_comparison <- comp %>%
  select(model,waic, se_waic)%>%
  left_join(weights)%>%
  arrange(desc(weight))

model_comparison <- read_csv("../saves/confirmatory_model_comparison.csv")

#write_csv(model_comparison, "../saves/confirmatory_model_comparison.csv")
```


# Visualize full model

## Posterior distributions for fixed effects
```{r}
model_draws <- full_model%>% posterior_samples(pars = c("b_delay", "b_task_experienceyes", "b_norm_age", "b_cup_distance", "b_board_size", "b_trial"))%>%
    mutate(model= "baseline")%>%
  pivot_longer(names_to = "estimate", values_to = "value", cols = c(b_delay,b_task_experienceyes,b_norm_age, b_cup_distance, b_board_size, b_trial ))%>%
  mutate(estimate = str_remove(estimate,"b_"))

model_draws %>%
  ggplot(.,aes(x = value, y= estimate, fill = factor(stat(quantile))))+
  geom_vline(xintercept = 0, lty = 2)+
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = c(0.025, 0.975)
  ) +
  theme_minimal() +
  scale_fill_manual(
   values = c("firebrick", "NA", "firebrick")
  )+
  guides(fill = F)
```

```{r}
ggsave("../figures/confirmatory_analysis_main_effects.png")
```


