---
title: "ManyPrimates1_Confirmatory_Analysis"
output:
  html_notebook:
    code_folding: hide
    css: style.css
    theme: paper
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(ape)
library(tidyverse)
library(readxl)
library(googledrive)
library(brms)
library(broom)
library(ggthemes)
library(tidybayes)
library(ggridges)
library(ggpubr)
```


# NOTE
Running the models takes quite some time (2h on a HPC) - outputs are too large to put on GitHub (~600MB). If you want to access the model outputs directly, you have to download them via the following link. Code that follows expects files to be in `saves/`.

```{r}
# download zip file from Max Planck cloud storage
options(timeout=3600) # 1h, maybe you need more depending on your internet speed
download.file("https://keeper.mpdl.mpg.de/f/3de147c222e946e78f59/?dl=1", 
              destfile = "../saves/saves.zip")
# unpack zip file to get .rds files
unzip("../saves/saves.zip", exdir = "../saves/")
```

If the code doesn't work, you can download the files here: `https://keeper.mpdl.mpg.de/f/3de147c222e946e78f59/?dl=1` but you will have to put them manually into the `saves` directory.

# Data processing

## Read in STM data file

Read in data, re-code condition and scale predictors. 

```{r}
data <- read_csv("../data/merged_data/ManyPrimates_mp1_merged_data.csv")%>%
  mutate(delay = if_else(condition == "short", -1,
                         if_else(condition == "medium", 0, 1)),
         trial = scale(trial, center = T, scale = T),
         norm_age = scale(norm_age, center = T, scale = T),
         cup_distance = scale(cup_distance, center = T, scale = T),
         board_size = scale(board_size, center = T, scale = T))
```

# Confirmatory analysis

## Full model

```{r}
# full_model <- brm(correct ~ delay +
#                     task_experience + norm_age + cup_distance + board_size + trial +
#                    (1 + delay + trial | subject_site + site + species),
#                    data=data,
#                    family=bernoulli(),
#                    chains = 4,
#                    iter= 5000,
#                    cores= 4,
#                    prior = prior(skew_normal(0, 1.5, 3), class=Intercept) +
#                            prior(skew_normal(-1.4, 2, 3), class=b, coef=delay) +
#                            prior(normal(0, 1), class=b, coef=board_size) +
#                            prior(normal(0, 1), class=b, coef=trial) +
#                            prior(normal(0, 1), class=b, coef=cup_distance) +
#                            prior(normal(0, 1), class=b, coef=norm_age) +
#                            prior(normal(0, 2), class=b, coef=task_experienceyes) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=subject_site) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=site) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=species) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=subject_site) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=site) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=species) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=subject_site) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=site) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=species) +
#                            prior("lkj(1)", class = "cor")
#                 )
# 
# summary(full_model)
# 
# full_model <- add_criterion(full_model, c("loo", "waic"))
# 
# full_model %>% saveRDS("../saves/conf_full_model.rds")

full_model <- readRDS("../saves/conf_full_model.rds")
```

## Null model

```{r}
# null_model <- brm(correct ~ task_experience + norm_age + cup_distance + board_size + trial +
#                    (1 + delay + trial | subject_site + site + species),
#                    data=data,
#                    family=bernoulli(),
#                    chains = 4,
#                    iter= 5000,
#                    cores= 4,
#                    prior = prior(skew_normal(0, 1.5, 3), class=Intercept) +
#                            #prior(skew_normal(-1.4, 2, 3), class=b, coef=delay) +
#                            prior(normal(0, 1), class=b, coef=board_size) +
#                            prior(normal(0, 1), class=b, coef=trial) +
#                            prior(normal(0, 1), class=b, coef=cup_distance) +
#                            prior(normal(0, 1), class=b, coef=norm_age) +
#                            prior(normal(0, 2), class=b, coef=task_experienceyes) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=subject_site) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=site) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=species) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=subject_site) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=site) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=species) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=subject_site) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=site) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=species) +
#                            prior("lkj(1)", class = "cor")
#                 )
# 
# summary(null_model)
# 
# null_model <- add_criterion(null_model, c("loo", "waic"))
# 
# null_model %>% saveRDS("../saves/conf_null_model.rds")

null_model <- readRDS("../saves/conf_null_model.rds")

```

# Model comparison

```{r}
# comp_waic <-  loo_compare(
#   full_model,
#   null_model,
#   criterion = c("waic"))%>%
#   as_tibble(rownames = "model")
# 
# comp_loo <-  loo_compare(
#   full_model,
#   null_model,
#   criterion = c("loo"))%>%
#   as_tibble(rownames = "model")
# 
# weights_waic <- model_weights(full_model, null_model, weights = "waic")%>%
#   as_tibble() %>%
#   rename(waic_weight = value) %>%
#   mutate(model  = c("full_model", "null_model"),
#          waic_weight = waic_weight %>% round(digits = 4)) %>%
#   select(model, waic_weight)
# 
# weights_loo <- model_weights(full_model, null_model, weights = "loo")%>%
#   as_tibble() %>%
#   rename(loo_weight = value) %>%
#   mutate(model  = c("full_model", "null_model"),
#          loo_weight = loo_weight %>% round(digits = 4)) %>%
#   select(model, loo_weight)
# 
# # Combine weights and scores
# model_comparison <- comp_waic %>%select(model,waic, se_waic)%>%
#   left_join(weights_waic)%>%
#   left_join(comp_loo %>%select(model,looic, se_looic))%>%
#   left_join(weights_loo)%>%
#   arrange(desc(waic_weight))
# 
# write_csv(model_comparison, "../saves/confirmatory_model_comparison.csv")

model_comparison <- read_csv("../saves/confirmatory_model_comparison.csv")

```

# Visualize full model

## Posterior distributions for fixed effects
```{r}
model_draws <- full_model%>% posterior_samples(pars = c("b_delay", "b_task_experienceyes", "b_norm_age", "b_cup_distance", "b_board_size", "b_trial"))%>%
    mutate(model= "baseline")%>%
  pivot_longer(names_to = "estimate", values_to = "value", cols = c(b_delay,b_task_experienceyes,b_norm_age, b_cup_distance, b_board_size, b_trial ))%>%
  mutate(estimate = str_remove(estimate,"b_"))

model_draws %>%
  ggplot(.,aes(x = value, y= estimate, fill = factor(stat(quantile))))+
  geom_vline(xintercept = 0, lty = 2)+
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = c(0.025, 0.975)
  ) +
  theme_minimal() +
  scale_fill_manual(
   values = c("firebrick", "NA", "firebrick")
  )+
  guides(fill = F)
```

```{r}
ggsave("../figures/confirmatory_analysis_main_effects.png")
```

# Exploratory analysis for age

## Interaction model

In the pilot, we included age in an interaction with delay. We add this model here as part of the exploratory analysis. 

```{r}
# int_model <- brm(correct ~ delay * norm_age +
#                     task_experience + cup_distance + board_size + trial +
#                    (1 + delay + trial | subject_site + site + species),
#                    data=data,
#                    family=bernoulli(),
#                    chains = 4,
#                    iter= 5000,
#                    cores= 4,
#                    prior = prior(skew_normal(0, 1.5, 3), class=Intercept) +
#                            prior(skew_normal(-1.4, 2, 3), class=b, coef=delay) +
#                            prior(normal(0, 1), class=b, coef=board_size) +
#                            prior(normal(0, 1), class=b, coef=trial) +
#                            prior(normal(0, 1), class=b, coef=cup_distance) +
#                            prior(normal(0, 1), class=b, coef=norm_age) +
#                            prior(normal(0, 1), class=b, coef=`delay:norm_age`) +
#                            prior(normal(0, 2), class=b, coef=task_experienceyes) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=subject_site) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=site) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=species) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=subject_site) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=site) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=species) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=subject_site) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=site) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=species) +
#                            prior("lkj(1)", class = "cor")
#                 )
# 
# summary(int_model)
# 
# int_model <- add_criterion(int_model, c("loo", "waic"))
# 
# int_model %>% saveRDS("../saves/conf_int_model.rds")

int_model <- readRDS("../saves/conf_int_model.rds")

```

## Quadratic model

We also add a model that assumes a non-linear relation between age and performance. 

```{r}
# quad_model <- brm(correct ~ delay + norm_age + I(norm_age^2) +
#                     task_experience + cup_distance + board_size + trial +
#                    (1 + delay + trial | subject_site + site + species),
#                    data=data,
#                    family=bernoulli(),
#                    chains = 4,
#                    iter= 5000,
#                    cores= 4,
#                    prior = prior(skew_normal(0, 1.5, 3), class=Intercept) +
#                            prior(skew_normal(-1.4, 2, 3), class=b, coef=delay) +
#                            prior(normal(0, 1), class=b, coef=board_size) +
#                            prior(normal(0, 1), class=b, coef=trial) +
#                            prior(normal(0, 1), class=b, coef=cup_distance) +
#                            prior(normal(0, 1), class=b, coef=norm_age) +
#                            prior(normal(0, 1), class=b, coef=Inorm_ageE2) +
#                            prior(normal(0, 2), class=b, coef=task_experienceyes) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=subject_site) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=site) +
#                            prior(normal(0, 1), class=sd, coef=Intercept, group=species) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=subject_site) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=site) +
#                            prior(normal(0, 0.5), class=sd, coef=delay, group=species) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=subject_site) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=site) +
#                            prior(normal(0, 0.5), class=sd, coef=trial, group=species) +
#                            prior("lkj(1)", class = "cor")
#                 )
# 
# summary(quad_model)
# 
# quad_model <- add_criterion(quad_model, c("loo", "waic"))
# 
# quad_model %>% saveRDS("../saves/conf_quad_model.rds")

quad_model <- readRDS("../saves/conf_quad_model.rds")
```

## Model comparison

```{r}
# comp_waic_exp <-  loo_compare(
#   full_model,
#   int_model,
#   quad_model,
#   criterion = c("waic"))%>%
#   as_tibble(rownames = "model")
# 
# comp_loo_exp <-  loo_compare(
#   full_model,
#   int_model,
#   quad_model,
#   criterion = c("loo"))%>%
#   as_tibble(rownames = "model")
# 
# weights_waic_exp <- model_weights(full_model, int_model,quad_model, weights = "waic")%>%
#   as_tibble() %>%
#   rename(waic_weight = value) %>%
#   mutate(model  = c("full_model", "int_model","quad_model"),
#          waic_weight = waic_weight %>% round(digits = 4)) %>%
#   select(model, waic_weight)
# 
# weights_loo_exp <- model_weights(full_model, int_model, quad_model, weights = "loo")%>%
#   as_tibble() %>%
#   rename(loo_weight = value) %>%
#   mutate(model  = c("full_model","int_model", "quad_model"),
#          loo_weight = loo_weight %>% round(digits = 4)) %>%
#   select(model, loo_weight)
# 
# # Combine weights and scores
# model_comparison_exp <- comp_waic_exp %>%select(model,waic, se_waic)%>%
#   left_join(weights_waic_exp)%>%
#   left_join(comp_loo_exp %>%select(model,looic, se_looic))%>%
#   left_join(weights_loo_exp)%>%
#   arrange(desc(waic_weight))
# 
# write_csv(model_comparison_exp, "../saves/confirmatory_model_comparison_exploratory.csv")

model_comparison_exp <- read_csv("../saves/confirmatory_model_comparison_exploratory.csv")

```
